Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-17 nips_2018.tex
Tweaking figure - feedback: is it too much?
big table at the beginning

Imagine that you are asked to write a program to edit some text, and
told that you should change the text ``Nancy FreeHafer'' to
``Dr. Nancy''.  From this example, you likely infer that ``Jane
Goodall'' should be ``Dr. Jane'', drawing upon your prior knowledge of
text, like that words are separated by spaces. For few-shot program learning problems
like these, the program synthesis and AI communities have had great
success engineering program learning systems,
not just for text editing~\cite{gulwani2011automating} but also for graphics
programs~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon having a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to write programs in that learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_2$\code{ s " "))}\\\cmidrule{2-2}
  \begin{tabular}{c}
    \sc{Library}\\
    \sc{(DSL)}
    \end{tabular}&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a \textbf{library} of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which library components are
useful for which kinds of tasks.
Together, the library and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{figure}
  \begin{tabular}{lcl}
    \toprule
    Domain&Example Task&Part of the learned DSL\\\midrule
    Lists&
    \begin{tabular}{rcl}
      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
    \end{tabular}
    &
    \begin{tabular}{l}
      \code{(foldr nil (lambda (a b) (cons a b)))}\\
          \hspace{0.5cm} \emph{(appends lists)}
      \end{tabular}

    \\    Strings&
    \begin{tabular}{rcl}
      Temple Anna H& $\longrightarrow$& TAH\\
      Lara Gregori&$\longrightarrow$&LG
      \end{tabular}
    &
    \begin{tabular}{l}
      \code{(map (lambda (x) (if (= x a) b x)))}\\
      \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
      \end{tabular}

@@

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We were inspired by EC,
and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
neurally guided program search.
In the experiments section,
we compared directly with EC,
showing empirical improvements.
The contribution of this work is
then
an effective algorithm for learning DSLs which also trains a neural net to
search for programs in the DSL.

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} step uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} steps update the DSL and the
recognition model, respectively.  Crucially, these steps synergistically bootstrap 
off each other:
\begin{comment}
  \begin{wrapfigure}{r}{0.5\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.\footnote{For example, for string editing,
  the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a generative model over programs.
We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x,p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x,p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

Together with its weight vector,
a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
In the supplement, we define this distribution  
by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. Care must be taken to ensure that programs are well-typed
and that variable scoping rules are obeyed.
With this distribution in hand, we search for programs by enumerating
$\lambda$-calculus expressions in decreasing order of their probability under
$(\mathcal{D},\theta)$.

@@

  A drawback of using an enumerative search algorithm is that we have no
efficient means of solving for arbitrary constants that might occur in the
program. In Sec.~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
%% In Sec.~\ref{textSection}
%% we will show that the bottom-up neural recognition model can learn
%% which discrete constants should be included in a program.

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting subexpressions from
programs already in the frontiers.
These subexpressions
are fragments of the original programs, and can introduce new variables (Fig.~\ref{fragmentExample}),
which then become new functions in the DSL.
The idea of storing and reusing
fragments of expressions comes from Fragment Grammars~\cite{tim} and Tree-Substitution Grammars~\cite{cohn2010inducing}.

@@

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
An obstacle here, however,
is that
our enumerative search procedure
has no means of generating string constants,
and so we incorporate  string-valued parameters as a primitive,
  defining $\probability[x|p]$ by marginalizing out the
  values of the string via dynamic programming.
  In  Section~\ref{regressionSection},
  we will use a similar trick to synthesize programs containing real numbers using gradient descent.

We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 32/108 of the problems with an average search time of 11 minutes.
After learning,
it solves 80/108, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 86/108 problems,
and does so with a
\emph{different hand-engineered DSL for each problem}.
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

>>>>

Imagine that you are asked to write a program to edit some text, and
told that you should change the text ``Nancy FreeHafer'' to
``Dr. Nancy''.  From this example, you likely infer that ``Jane
Goodall'' goes to ``Dr. Jane'', drawing upon your prior knowledge of
text, like that words are separated by spaces. For few-shot program learning problems
like these, the program synthesis and AI communities have had many successes,
not just for text editing~\cite{gulwani2011automating} but also for graphics
programs~\cite{ellis2017learning}, planning
procedures~\cite{devlin2017neural}, and many others.  However, the
success of these systems hinges upon having a carefully
hand-engineered \textbf{Domain Specific Language (DSL)}.  DSLs
impart prior knowledge of a domain
by providing a restricted set
of finely-tuned programming primitives: for
%that are finely tuned to the problems in the domain: for
text editing,
these are primitives like appending and splitting on characters.
In this work,
we consider the problem of
building agents that solve program learning tasks,
and also the problem of acquiring
the prior knowledge necessary to quickly
solve these tasks (Figure~\ref{firstPageFigure}).
Our solution is an algorithm that learns a DSL
while jointly training a neural network to help write programs in the learned DSL.
\begin{figure}[h]
  \begin{minipage}[c]{0.68\textwidth}
\hspace{-0.5cm}      \begin{tabular}{r|c|}\cmidrule{2-2}
  \sc{Task}&
  \begin{tabular}{rcl}
  Nancy FreeHafer&$\longrightarrow$& Dr. Nancy\\
% Andrew Cencici&$\longrightarrow$&Dr. Andrew\\
 Jane Goodall&$\longrightarrow$&???
  \end{tabular}\\\cmidrule{2-2}
  \sc{Program}&$f(\text{\code{s}})\,=\,$\code{(}$f_0$\code{ "Dr." (}$f_1$\code{ s ' '))}\\\cmidrule{2-2}
    \sc{DSL}
&\begin{tabular}{l}
    $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b ($\lambda$ (x y)}\\
    \phantom{$f_0($\code{a}$,$\code{b}$) \,=\, $\code{(}}\code{(cons x y)))}\\
    \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
    $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s ($\lambda$ (x a)}\\
    \phantom{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{  }}\code{(if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
    \end{tabular}\\\cmidrule{2-2}
  \end{tabular}
  \end{minipage}%
  \begin{minipage}[c]{0.31\textwidth}
      \caption{\textbf{Task}: Few-shot learning problem. Model solves tasks by writing \textbf{program}s, and jointly learns a library of reusable subroutines that are shared across multiple tasks, called a \textbf{Domain Specific Language (DSL)}. Program writing is guided by a neural network trained jointly with the library.}\label{firstPageFigure}
    \end{minipage}
\end{figure}

%% We think of
%% solutions to these tasks as being well represented by programs,
%% and so our problem can be stated as follows:
%% how should an agent learn to write programs?
We take inspiration from two sources:
(1) Good software engineers compose libraries
of reusable subroutines
that are shared across
related programming tasks.
Returning to Figure~\ref{firstPageFigure},
a good string processing library should support appending strings and splitting on spaces --
exactly the prior knowledge needed to solve the task in Figure~\ref{firstPageFigure}.
(2) Skilled human programmers
can quickly recognize what kinds of programming idioms and library routines would be useful for solving
the task at hand, even if they cannot instantly work out the details.
We combine these two ideas into
an algorithm called \systemEnding,
which takes as input a collection of
programming \textbf{tasks},
and then jointly
solves three problems:
(1) Writing programs that solve the tasks;
(2) Composing a library (DSL) of
domain-specific
subroutines -- which allow the agent to
more compactly write programs in the domain,
and (3) Training a neural network
to recognize
which DSL components are
useful for which kinds of tasks.
Together, the DSL and neural net
encode the domain specific knowledge needed to
quickly write programs.

@@

We apply \system to four domains:
list processing; FlashFill-style~\cite{gulwani2011automating} string editing; and symbolic regression.
 For each of these we initially provide a generic
 set of programming primitives.
Our algorithm then discovers
its own DSL for expressing solutions in the domain (Tbl.~\ref{initialExampleDSL}).
\begin{figure}
  \begin{tabular}{lcl}
    \toprule
    Domain&Example Task&Part of the learned DSL\\\midrule
    Lists&
    \begin{tabular}{rcl}
      \code{[7\, 0\, 2]}&$\longrightarrow$&\code{[7\, 0\, 2\, 4]} \\
      \code{[3\, 9]}&$\longrightarrow$&\code{[3\, 9\, 4]}
    \end{tabular}
    &
    \begin{tabular}{l}
      \code{(foldr nil ($\lambda$ (a b) (cons a b)))}\\
          \hspace{0.5cm} \emph{(appends lists)}
      \end{tabular}

    \\    Strings&
    \begin{tabular}{rcl}
      Temple Anna H& $\longrightarrow$& TAH\\
      Lara Gregori&$\longrightarrow$&LG
      \end{tabular}
    &
    \begin{tabular}{l}
      \code{(map ($\lambda$ (x) (if (= x a) b x)))}\\
      \hspace{0.5cm} \emph{(replace occurrences of a w/ b)}\\
      \end{tabular}

@@

\begin{tabular}{ccc}\toprule
  \begin{tabular}{c}
    \code{[7\, 0\, 2]}$\to $\code{[2\, 0\, 7]} \\
      \code{[2\, 4\, 9\, 3]}$\to $\code{[3\, 9\, 4\, 2]}\\
      \code{$f(x) = $}\code{($f_0$ nil)}\\\\
      \code{[99\, 4\, 2]}$\to $\code{99}\\
      \code{[2\, 1\, 4\, 6]}$\to $\code{6}\\
      \code{[8\, 7\, 7\, 7]}$\to $\code{8}\\
      \code{($f_1$ )}
  \end{tabular}&
  \begin{tabular}{c}
    +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
    +83 973-831$\to $83.973.831\\
    $f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,} ($f_0$ \codechar{.} \codechar{ } (cdr s)))}    \\\\
Temple Anna H $\to $TAH\\
Lara Gregori$\to $LG\\
$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}\\

  \end{tabular}&

  \begin{tabular}{c}
    \includegraphics[width = 1cm]{figures/functions/4.png}\\
    \code{$f($x$) = $($f_1$ -0.8 1.6)}\\
    \includegraphics[width = 1cm]{figures/functions/112.png}\\
    \code{$f($x$) = $($f_0$ -0.8 1.6)}\\

    \end{tabular}

  \\\midrule&

  \begin{tabular}{l}
          $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x) (if (= x a) b x)) s)}\\
      \hspace{0.5cm}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car ($\lambda$ (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
    \end{tabular}&

  \begin{tabular}{l}
    $f_0($\code{x}$)\,=\,$\code{(+ x real)}\quad
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
    \end{tabular}

      \\\bottomrule

  \end{tabular}

Prior work on program learning has largely assumed a fixed, hand-engineered DSL,
both in classic symbolic program learning approaches (e.g., Metagol:~\cite{muggleton2015meta},
FlashFill:~\cite{gulwani2011automating}),
neural approaches  (e.g., RobustFill:~\cite{devlin2017robustfill}), and hybrids of neural and
symbolic methods (e.g., Neural-guided deductive search:~\cite{ngds}, DeepCoder:~\cite{balog2016deepcoder}).
A notable exception is the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316},
which also learns a library of subroutines.
We were inspired by EC,
and go beyond it by giving a new algorithm for learning DSLs, as well as a way of combining DSL learning with
neurally guided program search.
In the experiments section,
we compared directly with EC,
showing empirical improvements.
The contribution of this work is
then
an effective algorithm for learning DSLs which jointly trains a neural net to
search for programs in the DSL.

@@

 Our goal is to induce a DSL while finding programs solving each of the
tasks.  We take inspiration primarily from the
Exploration-Compression algorithm for bootstrap
learning~\cite{Dechter:2013:BLV:2540128.2540316}.  Exploration-Compression
alternates between exploring the space of solutions to a set of tasks,
and compressing those solutions to suggest new search primitives
for the next exploration stage.
We extend these ideas into an inference strategy that iterates
through three steps: an \textbf{Explore} step uses the current DSL and
recognition model to search for programs that solve the tasks.  The
\textbf{Compress} and \textbf{Compile} steps update the DSL and the
recognition model, respectively.  Crucially, these steps bootstrap 
each other:
\begin{comment}
  \begin{wrapfigure}{r}{0.5\textwidth}
  \begin{tikzpicture}
    \begin{scope}[shift = {(1,-1)}]
    \node[align = center](synthesis) at (6,4) {Search for \\programs};
    \node[align = center](DSL) at (3,1) {DSL};
    \node[align = center](recognitionModel) at (9,1) {Recognition \\model};

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.%% \footnote{For example, for string editing,
%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
%% and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a generative model over programs.
We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t. the frontiers (Exploration) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t. the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t. the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\textbf{Neural recognition model: tractably maxing $\lowerBound$ w.r.t. the
  frontiers.}  Here we train %% a neural network, $q$, to predict a
%% distribution over programs conditioned on a task. The objective of $q$
%% is
$q(p|x)$ to assign high probability to programs $p$ where
$\probability[x|p]\probability[p|\mathcal{D},\theta]$ is large, because including those programs
in the frontiers will most increase $\lowerBound$.  %% With $q$ in hand we can find programs for
%% the frontier $\mathcal{F}_x$ by searching for programs that maximize
%% $q(p|x)$.
%% Rather than directly predicting a distribution over $p$ conditioned on $x$,
%% the recognition model predicts a distribution $\theta^{(x)}$ over components of the DSL.
%% This taps into the intuition that programming is primarily a top-down activity:
%% as human programmers, we often first decide what kind of
%% programming constructs we might need to use,
%% and then we figure out how to assemble them in to the desired program.
%% This approach implements an amortized
%% inference scheme~\cite{le2016inference,ritchie2016deep} for the generative model in
%% Fig.~\ref{graphicalModel}.

@@

Together with its weight vector,
a DSL defines a distribution over programs, $\probability[p|\mathcal{D},\theta]$.
In the supplement, we define this distribution  
by specifying a procedure for drawing samples from $\probability[p|\mathcal{D},\theta]$. %%  Care must be taken to ensure that programs are well-typed
%% and that variable scoping rules are obeyed.
%% With this distribution in hand, we search for programs by enumerating
%% $\lambda$-calculus expressions in decreasing order of their probability under
%% $(\mathcal{D},\theta)$.

@@

  However, a drawback of   enumerative search  is that we have no
efficient means of solving for arbitrary constants that might occur in a
program. In Sec.~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
%% In Sec.~\ref{textSection}
%% we will show that the bottom-up neural recognition model can learn
%% which discrete constants should be included in a program.

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{equation}
      \log \probability[\mathcal{D}] + \argmax_{\theta}\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta] + \log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \label{AIC}
  \end{equation}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Fig.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

@@

\subsection{String Editing}\label{textSection}
Synthesizing programs that edit strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other systems successful.

Because our enumerative search procedure has no means of generating
string constants, we have the enumerator propose programs with string-valued holes --
e.g., in Fig.~\ref{firstPageFigure},
we enumerate $\text{\code{(}}f_0\code{ string (}f_2\text{\code{ s ' '))}}$ --
and define $\probability[x|p]$ by marginalizing out the values of the
strings via dynamic programming.  In Section~\ref{regressionSection},
we will use a similar trick to synthesize programs containing real
numbers, but using gradient descent instead of dynamic programming.

We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our system on problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems with an average search time of 235 seconds.
After learning,
it solves 74.1\%, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 79.6\% problems.
But, SyGuS comes with a
different hand-engineered DSL \emph{for each text editing problem}.
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.
